Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-07-03 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
use l2 norm values, not L2

The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.  Error is largest on the BTF grid with \(\ell^2 = \num{0.00791}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).  The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.

The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.0210} to \num{0.0318}.  The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.00791} to \num{0.00767}.

@@

% TABLES
\begin{table*}
	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.  \(\ell^2\) error norms were not calculated by this code.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
\hline
\end{tabular}
\end{table*}

>>>>

The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.  Error is largest on the BTF grid with \(\ell^2 = \num{0.107}\) but is significantly reduced on the SLEVE grid with \(\ell^2 = \num{0.0146}\).  The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.

The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.  The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.107} to \num{0.104}.

@@

% TABLES
\begin{table*}
	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, \(\ell^2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                  & Fourth order        & \(\ell^2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.107   & 0.0146          & 0.00784        & 0.00784     \\
                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.0341  & 0.235           & 184            & {---}        \\
		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
\hline
\end{tabular}
\end{table*}
